R-LIME: LIME法における矩形制約と最適化

### 2024-03-18 paper/paper.tex
Update submodule myref
ICPR 3rd version

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb} % for \mathbb
\usepackage{array} % for m column type
\usepackage{bbm} % for \mathbbm
\usepackage{booktabs} % for \toprule, \midrule, \bottomrule
\usepackage{cleveref} % for \cref
\usepackage{csvsimple} % for \csvreader
\usepackage[T1]{fontenc} % for T1 font encoding
\usepackage{graphicx} % for \includegraphics
\usepackage{pgffor} % for \foreach
\usepackage[labelformat=simple]{subcaption} % for subfigure environment

@@

\ifnum\mode=0
  \pagestyle{plain}
  \pagenumbering{roman}
  \tableofcontents
  \newpage
  \listoffigures\listoftables\clearpage
  \AddToHook{cmd/section/before}{\clearpage}
  \pagestyle{plain}
  \pagenumbering{arabic}
\else

  % The abstract should briefly summarize the contents of the paper in
  % 150--250 words.
  \begin{abstract}
    In recent years,
    complex machine learning models such as deep neural networks
    have been used in various industrial fields due to their high accuracy.
    However,
    its complexity has been a major obstacle to implementation
    in decision-making situations where transparency of the decision process
    is required.
    In order to address this problem,
    various post-hoc explanation methods have been proposed,
    but they have not been able to achieve interpretability of
    both the explanation and its scope.
    We propose a new method, R-LIME,
    which interprets complex classifiers in an interpretable scope.
    R-LIME locally approximizes a complex decision boundary linearly
    in a rectangular region
    and maximizes the region as long as the accuracy of the linear
    classifier is higher than a specified threshold.
    The resulting rectangular region is interpretable for users because it is
    expressed as a conjunction of feature predicates.
    We demonstrate the effectiveness of the proposed method through qualitative
    and quantitative experiments using a real-world dataset.
    Finally, we discuss limitations and future work of the proposed method.
    % 148 words
    \keywords{Interpretable machine learning \and Local surrogate model}
  \end{abstract}
\fi

\section{Introduction}
Machine learning models, such as deep learning and random forests,
have been widely employed in various industrial applications
due to their significant improvement in accuracy in recent years.
However,
the increasing complexity and black-box nature of these models pose challenges,
particularly in critical decision-making scenarios like healthcare and finance,
where the opacity of decision process becomes a major obstacle to
implementation.
Consequently,
there has been extensive research in the field of post-hoc explanations
for machine learning models~\cite{%
  guidotti2018local,radulovic2023bella,ribeiro2016why,ribeiro2018anchors}.
Existing post-hoc explanation methods are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure.
Furthermore, model-agnostic methods are classified into
\emph{global} and \emph{local} methods based on their locality in input space.

This paper specifically focuses
on \emph{model-agnostic} and \emph{local} methods,
which do not rely on the models' structure
and explain the models' behavior
in the vicinity of a given input (focal point).
For example, LIME (Local Interpretable Model-agnostic Explanations)
\cite{ribeiro2016why} approximates the complex model's decision boundary
locally using a simpler model around the given focal point.
Although LIME enables users to interpret the local behavior of complex models,
it does not explicitly provide the effective scope of the explanation
\cite{ribeiro2018anchors}.
This limitation can lead
users to overgeneralize the knowledge derived from the explanation.
Without a clear indication of the scope of the explanation,
users might mistakenly apply
the derived knowledge to instances outside this scope,
potentially leading to misunderstanding of the black-box model's behavior.

To address this limitation of LIME,
we propose a new method called R-LIME (Ruled LIME),
which interprets complex classifiers in an interpretable region.
R-LIME locally approximates a complex decision boundary linearly
in a rectangular region and maximizes the region
as long as the accuracy of the linear classifier is
higher than a given threshold.
The resulting rectangular region is interpretable for users because it is
expressed as a conjunction of feature predicates (\cref{fig:example-rlime}).

\section{Related Work}
In this chapter,
we overview existing research on post-hoc explanations,
which explain the behavior of complex models already trained.
They are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure.

Model-dependent methods,
such as Deep Taylor Decomposition (DTD)~\cite{montavon2017explaining}
and Layer-wise Relevance Propagation (LRP)~\cite{bach2015pixel},
primarily focus on deep neural networks and
explain the models' behavior using their parameters~\cite{samek2021explaining}.
While these methods can provide detailed explanations
(e.g., layer-wise explanations for neural networks)
using the models' internal parameters,
it is often challenging
to apply the same method to models with different structures.

On the other hand,
model-agnostic methods
do not rely on the models' parameters and utilize only their output.
Although these methods are applicable to any model,
they cannot explain the reasoning process inside the model.
Furthermore, model-agnostic methods are classified into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as Partial Dependence Plots (PDP)~\cite{friedman2001greedy}
and Accumulated Local Effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space,
seeking explanations valid for any input.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as LIME (Local Interpretable Model-agnostic Explanations)~\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors},
% SHAP~\cite{lundberg2017unified},
% EXPLAN~\cite{rasouli2020explan},
% LORE (Local Rule-based Explanations)~\cite{guidotti2018local}
% and BELLA (Black box model Explanations by Local Linear Approximations)
% \cite{radulovic2023bella},
explain output of the model for a specific input
(focal point) in the vicinity of that input.
While local methods offer more simple and accurate explanations
than global methods,
the scope of the explanation is limited to the vicinity of the focal point.

\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \ifnum\mode=0
      \vspace{-1.79cm}
    \else
      \vspace{-2.24cm}
    \fi
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method).
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}
In this paper,
we focus on local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME~\cite{ribeiro2016why} approximates the complex decision boundary
locally using a linear model around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) provides suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate the effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For example,
user may apply the insights derived from the explanation
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is incorrect obviously.
Anchor~\cite{ribeiro2018anchors} represents a rectangular region
containing the focal point,
that maximizes the probability of the black-box classifier outputting
the same label as the focal point.
While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility~\cite{ribeiro2018anchors}.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that changing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

% BELLA\cite{radulovic2023bella} addresses the non-optimization of the approximation region,
% a drawback of LIME,
% by learning a local linear classifier $g$
% using a subset $Z'$ of the dataset $Z$
% instead of perturbed samples.
% The process involves:
% \begin{enumerate}
% 	\item Computing the distance $d_x(z)$ for all instances $z$ in $Z$
% 	      with respect to the focal point $x$.
% 	\item Selecting the $k$ instances with the smallest $d_x(z)$,
% 	      forming a subset $Z_s(k) \subseteq Z$.
% 	\item Learning the local linear classifier $g$ using $Z_s(k)$
% 	      and calculating the similarity measure $\mathcal{R}(k)$ between $f$ and $g$
% 	      over $k=1,2,\dots,|Z|$.
% 	\item Obtaining the optimal $k^*$ by maximizing $\mathcal{R}(k)$.
% \end{enumerate}
% 
% BELLA aims to enhance the reliability of explanations
% by optimizing the approximation region for high-accuracy linear approximations.
% However, it assumes direct access to the dataset $Z$ used for training the classifier,
% making it inapplicable in scenarios where direct access to the dataset is restricted.
% Additionally,
% as BELLA presents the approximation region as a subset of the dataset,
% its interpretability is lower compared to Anchor.

\section{Proposed Method}
\begin{figure}[tbp]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-lime}
    \caption{%
      LIME: Locally approximates the decision boundary around the focal point.
    }\label{fig:lime}
  \end{subfigure}%
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-anchor}
    \caption{%
      Anchor: Maximizes coverage of a rectangular region
      containing the focal point under accuracy constraints.
    }\label{fig:anchor}
  \end{subfigure}
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-rlime3}
    \caption{%
      R-LIME:
      Maximizes coverage of a rectangular region
      under lower constraints on the accuracy of the linear classifier.
    }\label{fig:rlime}
  \end{subfigure}
  \caption[Visual comparison of LIME, Anchor and R-LIME]{%
    Visual comparison of LIME, Anchor and R-LIME (our proposed method).
    The solid line represents the rectangular region containing the focal
    point, and the dashed line represents the learned approximation model.
  }
\end{figure}
\subsection{Previous Work}
We specifically focuses on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.
\subsubsection[LIME]{%
  LIME (Local Interpretable Model-agnostic Explanations)
  \cite{ribeiro2016why}
}
LIME locally
approximates a trained black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The process involves:
\begin{enumerate}
  \item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \label{eq:lime_loss}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}
LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, due to not explicitly indicating the perturbation region,
users cannot assess the effective scope of the explanation
\cite{ribeiro2018anchors}.

@@

  \foreach \alg in {2,3,4}{%
      \begin{algorithm}[p]
        \small
        \input{tex/alg\alg}
      \end{algorithm}
    }
}

The algorithm of our method
is mainly based on that used in Anchor\cite{ribeiro2018anchors}.
For non-convex optimization problems like \cref{eq:main_problem},
greedy search are often used.
However,
greedy methods tend to converge to local optima, and to address this,
R-LIME utilizes beam search,
which selects multiple candidates at each iteration.
The pseudo code is shown in \cref{alg:greedy_search}.

@@

\subsubsection{Selection of Candidate Rules with Maximum Accuracy}
Given the set of generated candidate rules $\cands$,
the algorithm selects the top $B$ candidates with the highest accuracy.
This is treated as an optimal arm identification problem in the multi-armed bandit framework.
Each candidate rule $A_i\in\cands$ is considered an arm,
and their accuracy $\Prec(A_i)$ is treated as the distribution of rewards.
By sampling $z\sim\mathcal{D}(\cdot|A_i)$
and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial,
the algorithm updates $g_i$ using the sampled perturbation vector $z$
and the pseudo-label $f(z)$ after each trial.
To efficiently select the rule (arm) with the highest accuracy,
we employ the KL-LUCB algorithm \cite{kaufmann2013information}.
The pseudo code is shown in \cref{alg:best_cands}.
For tolerance $\epsilon\in[0,1]$, the KL-LUCB algorithm guarantees below:
\begin{equation}
  P(\underset{A\in\cands}{\min}\Prec(A)\ge
  \underset{A'\in\mathcal{A}}{\min}\Prec(A')-\epsilon)\ge1-\delta.
\end{equation}
However,
the KL-LUCB algorithm assumes that the reward distribution for each arm
remains unchanged,
while our method updates the classifier $g_i$ with each sampling,
which may not satisfy this assumption.
This issue is discussed further in \cref{sec:reward}.

@@

\ifnum\mode=1
  \subsection{Computational Complexity}
  Post-hoc explanation methods including LIME, Anchor, and R-LIME
  need to query the black-box model multiple times
  to get labels for perturbed samples, which is computationally expensive.
  The number of queries required for LIME is $|\mathcal{Z}_p|$,
  where $|\mathcal{Z}_p|$ is the number of perturbed samples.
  On the other hand,
  the number of queries required for Anchor and R-LIME is
  $\mathcal{O}(B\cdot m^2+m^2\cdot\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]})$,
  where $B$ is the beam width, $m$ is the number of features, and
  $\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ is the computational complexity of
  pure exploration problem of finding the $B$ best arms among $B\cdot m$ arms.
  $\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ depends on hyperparameters
  such as threshold $\tau$, confidence level $1-\delta$ and tolerance $\epsilon$.
\fi

@@

The dataset, after removing missing values,
was split into training data (7,639 instances) and test data (955 instances).
A random forest model with 50 trees was trained using the training data.
Subsequently, LIME and R-LIME explanations were generated
for two instances extracted from the test data (\cref{fig:instance}).
In R-LIME, logistic regression was used as the linear approximation model,
and the distribution $\mathcal{D}$ was a multivariate normal distribution
estimated from the training data.
The beam width was set to $B=10$,
the confidence coefficient to $1-\delta=0.95$,
and the tolerance of the KL-LUCB algorithm to $\epsilon=0.05$.
Accuracy thresholds $\tau$ were set to $\tau=0.70,0.80,0.90$.

@@

\subsection{Changes in Reward Distribution in Best Arm Identification}\label{sec:reward}
{%
  \renewcommand{\arraystretch}{1.1}
  \begin{table}[tbp]
    \centering
    \caption[Deviation between the estimated accuracy and the true accuracy]{%
      Deviation between the estimated accuracy and the true accuracy.
      Deviation was relatively small considering confidence level $1-\delta=0.95$.
    }\label{tab:reward}
    \input{tex/tab_exp_adv}
  \end{table}
}
In R-LIME,
the problem of selecting the rule with the highest accuracy is formulated
as the optimal arm identification problem in multi-armed bandit theory,
solved using the KL-LUCB algorithm \cite{kaufmann2013information}.
However, this algorithm assumes that the reward distribution remains constant,
while in R-LIME,
the reward distribution (accuracy of the linear approximation)
changes with every update of the approximation model after sampling.
Therefore, rewards obtained at an early stage
might influence the estimated value and deviate from the true value.

>>>>

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb} % for \mathbb
\usepackage{array} % for m column type
\usepackage{bbm} % for \mathbbm
\usepackage{booktabs} % for \toprule, \midrule, \bottomrule
\usepackage{cleveref} % for \cref
\usepackage{csvsimple} % for \csvreader
\usepackage[T1]{fontenc} % for T1 font encoding
\usepackage{graphicx} % for \includegraphics
\usepackage{pgffor} % for \foreach
\usepackage[labelformat=simple]{subcaption} % for subfigure environment
\usepackage{tikz}
\usetikzlibrary{positioning}

@@

\ifnum\mode=0
  \pagestyle{plain}
  \pagenumbering{roman}
  \tableofcontents
  \newpage
  \listoffigures\listoftables\clearpage
  \AddToHook{cmd/section/before}{\clearpage}
  \pagestyle{plain}
  \pagenumbering{arabic}
\else
  % The abstract should briefly summarize the contents of the paper in
  % 150--250 words.
  \begin{abstract}
    In recent years,
    complex machine learning models such as deep neural networks
    have been used in various industrial fields due to their high accuracy.
    However,
    its complexity has been a major obstacle to implementation
    in decision-making situations where transparency of the decision process
    is required.
    In order to address this problem,
    various post-hoc explanation methods have been proposed,
    but they have not been able to achieve interpretability of
    both the explanation and its scope.
    We propose a new method, R-LIME,
    which interprets complex classifiers in an interpretable scope.
    R-LIME locally approximizes a complex decision boundary linearly
    in a rectangular region
    and maximizes the region as long as the accuracy of the linear
    classifier is higher than a specified threshold.
    The resulting rectangular region is interpretable for users because it is
    expressed as a conjunction of feature predicates.
    We demonstrate the effectiveness of the proposed method through qualitative
    and quantitative experiments using a real-world dataset.
    Finally, we discuss limitations and future work of the proposed method.
    % 148 words
    \keywords{Interpretable machine learning \and Local surrogate model}
  \end{abstract}
\fi

\section{Introduction}
Machine learning models, such as deep learning and random forests,
have been widely employed in various industrial applications
due to their significant improvement in accuracy in recent years.
However,
the increasing complexity and black-box nature of these models pose challenges,
particularly in critical decision-making scenarios like healthcare and finance,
where the opacity of decision process becomes a major obstacle to
implementation.
Consequently,
there has been extensive research in the field of post-hoc explanations
for machine learning models~\cite{%
  guidotti2018local,radulovic2023bella,ribeiro2016why,ribeiro2018anchors}.
Existing post-hoc explanation methods are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure.
Furthermore, model-agnostic methods are classified into
\emph{global} and \emph{local} methods based on their locality in input space.
\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \ifnum\mode=0
      \vspace{-1.79cm}
    \else
      \vspace{-2.24cm}
    \fi
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method).
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}

In this paper,
we focus on local model-agnostic methods.
An example of explanations by LIME~\cite{ribeiro2016why} and
Anchor~\cite{ribeiro2018anchors}
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME approximates the complex decision boundary
linearly around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) provides suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate the effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For this example,
user may apply the derived insights
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is obviously incorrect.
Anchor represents a rectangular region
containing the focal point,
that maximizes the probability of the black-box classifier outputting
the same label as the focal point.
While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility~\cite{ribeiro2018anchors}.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that changing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

To address this limitation of LIME,
we propose a new method called R-LIME (Ruled LIME),
which interprets complex classifiers in an interpretable region.
R-LIME locally approximates a complex decision boundary linearly
in a rectangular region and maximizes the region
as long as the accuracy of the linear classifier is
higher than a given threshold.
The rectangular region is interpretable for users because it is
expressed as a conjunction of feature predicates (\cref{fig:example-rlime}).
An example of the explanation by R-LIME for a sentiment prediction model
is illustrated in \cref{fig:example-rlime}.
It is clear that users can apply the insights derived from the explanation
only to the sentences containing the word ``not''.

\section{Related Work}
\begin{figure}[tbp]
  \centering
  \def\w{2.2}
  \def\ww{4.4}
  \def\h{0.7}
  \def\hh{1.4}
  \def\hhh{2.1}
  \def\hhhh{2.8}
  \begin{tikzpicture}[scale=0.9,nodestyle/.style={rectangle,node font=\normalsize\itshape,auto}]
    \node[nodestyle] (post-hoc) at (0,0){post-hoc};
    \node[nodestyle] (dependent) at (-\w,-\hh){model-dependent};
    \node[nodestyle] (agnostic) at (\w,-\hh){\textbf{model-agnostic}};
    \node[nodestyle] (global) at (0,-\hhhh){global};
    \node[nodestyle] (local) at (\ww,-\hhhh){\textbf{local}};
    \draw(post-hoc) -- (0,-\h);
    \draw(dependent) -- (-\w,-\h) -- (\w,-\h) -- (agnostic);
    \draw(agnostic) -- (\w,-\hhh);
    \draw(global) -- (0,-\hhh)  -- (\ww,-\hhh) -- (local);
  \end{tikzpicture}
  \caption{%
    Categorization of post-hoc explanation methods.
    We focus on \emph{model-agnostic} and \emph{local} methods,
    which explain model's local behavior using only its output.
  }\label{fig:post-hoc}
\end{figure}
In this chapter,
we overview existing research on post-hoc explanation methods,
which explain the behavior of black-box models already trained.
As shown in \cref{fig:post-hoc},
post-hoc methods are classified into several categories.

First, they are categorized into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure.
Model-dependent methods,
such as Deep Taylor Decomposition (DTD)~\cite{montavon2017explaining}
and Layer-wise Relevance Propagation (LRP)~\cite{bach2015pixel},
most of which focus on neural networks and
explain the model's behavior using its parameters~\cite{samek2021explaining}.
While these methods provide detailed explanations
(e.g., layer-wise explanations for neural networks),
it is often challenging
to apply the same method to models with different structures.
On the other hand,
model-agnostic methods use only the output of the model.
Although they are applicable to any model,
they cannot explain the reasoning process inside the model.

Then, model-agnostic methods are classified into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as Partial Dependence Plots (PDP)~\cite{friedman2001greedy}
and Accumulated Local Effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space.
% seeking explanations valid for any input.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as Local Interpretable Model-agnostic Explanations (LIME)
\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors},
% SHAP~\cite{lundberg2017unified},
% EXPLAN~\cite{rasouli2020explan},
% LORE (Local Rule-based Explanations)~\cite{guidotti2018local}
% and BELLA (Black box model Explanations by Local Linear Approximations)
% \cite{radulovic2023bella},
explain model's behavior in the vicinity of a specific input.
While local methods offer explanations more simple and accurate
than global methods,
the scope of the explanation is limited locally.

\section{Proposed Method}
\begin{figure}[tbp]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-lime}
    \caption{%
      LIME: Locally approximates the decision boundary around the focal point.
    }\label{fig:lime}
  \end{subfigure}%
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-anchor}
    \caption{%
      Anchor: Maximizes coverage of a rectangular region
      containing the focal point under accuracy constraints.
    }\label{fig:anchor}
  \end{subfigure}
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-rlime3}
    \caption{%
      R-LIME:
      Maximizes coverage of a rectangular region
      under lower constraints on the accuracy of the linear classifier.
    }\label{fig:rlime}
  \end{subfigure}
  \caption[Visual comparison of LIME, Anchor and R-LIME]{%
    Visual comparison of LIME, Anchor and R-LIME (our proposed method).
    The solid line represents the rectangular region containing the focal
    point, and the dashed line represents the learned approximation model.
  }
\end{figure}

\subsection{Previous Work}
We specifically focuses on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

\subsubsection[LIME]{%
  LIME (Local Interpretable Model-agnostic Explanations)
  \cite{ribeiro2016why}
}
LIME locally
approximates a black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The approximation is performed by the following two steps:
\begin{enumerate}
  \item Generating a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
        (i) $x$ is converted into a binary vector
        $x'\in\{0,1\}^{m'}$,
        (ii) perturbed samples are generated by drawing non-zero elements
        from $x'$ uniformly at random,
        and (iii) the perturbed samples are converted back to the original space.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \label{eq:lime_loss}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}
LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, due to not explicitly indicating the perturbation region,
users cannot assess the effective scope of the explanation
\cite{ribeiro2018anchors}.

@@

  \foreach \alg in {2,3,4}{%
      \begin{algorithm}[p]
        \small
        \input{tex/alg\alg}
      \end{algorithm}
    }
}
The algorithm of R-LIME
is mainly based on that used in Anchor\cite{ribeiro2018anchors}.
For non-convex optimization problems like \cref{eq:main_problem},
greedy search are often used.
However,
greedy methods tend to converge to local optima, and to address this,
R-LIME utilizes beam search,
which selects multiple candidates at each iteration.
The pseudo code is shown in \cref{alg:greedy_search}.

@@

\subsubsection{Selection of Candidate Rules with Maximum Accuracy}
Given the set of candidate rules $\cands$,
the algorithm selects the top $B$ candidates with the highest accuracy.
This problem can be formulated
as best arm identification in the multi-armed bandit framework.
Each candidate rule $A_i\in\cands$ is considered as an arm,
and its accuracy $\Prec(A_i)$ is considered as the distribution of rewards.
By sampling $z\sim\mathcal{D}(\cdot|A_i)$
and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial,
the algorithm updates $g_i$ using the sampled perturbation vector $z$
and the pseudo-label $f(z)$ after each trial.
To efficiently select the rule (arm) with the highest accuracy,
we employ the KL-LUCB algorithm \cite{kaufmann2013information}.
The pseudo code is shown in \cref{alg:best_cands}.
For tolerance $\epsilon\in[0,1]$, the KL-LUCB algorithm guarantees below:
\begin{equation}
  P(\underset{A\in\cands}{\min}\Prec(A)\ge
  \underset{A'\in\mathcal{A}}{\min}\Prec(A')-\epsilon)\ge1-\delta.
\end{equation}
However,
the KL-LUCB algorithm assumes that the reward distribution for each arm
remains unchanged,
while our method updates the classifier $g_i$ with each sampling,
which may not satisfy this assumption.
This issue is discussed further in \cref{sec:reward}.

@@

\ifnum\mode=1
  \subsection{Computational Complexity}
  % \subsubsection{Number of Queries}
  Post-hoc explanation methods including LIME, Anchor, and R-LIME
  need to query the black-box model multiple times
  to get labels for perturbed samples, which is computationally expensive.
  The number of queries required for LIME is $|\mathcal{Z}_p|$,
  where $|\mathcal{Z}_p|$ is the number of perturbed samples selected by the user.
  On the other hand,
  the number of queries required for Anchor and R-LIME is
  $\mathcal{O}(B\cdot m^2+m^2\cdot\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]})$,
  where $B$ is the beam width, $m$ is the number of features, and
  $\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ is the computational complexity of
  best arm identification finding the $B$ best arms among $B\cdot m$ arms.
  $\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$ depends on hyperparameters
  such as beam width $B$, confidence level $1-\delta$ and tolerance $\epsilon$.
  % \subsubsection{Updating the Linear Classifier}
  % For Anchor, each trial is composed of sampling a perturbation vector $z$
  % and getting output $f(z)$.
  % But for R-LIME, a trial
  % includes additionally updating the linear classifier $g$ by $z$ and $f(z)$.
  % If we use support vector machine (SVM) as the linear classifier and
  % update it by stochastic gradient descent (SGD)~\cite{robbins1951stochastic},
  % the computational complexity of updating $g$ for each trial is $\mathcal{O}(m^2)$.
\fi

@@

The dataset, after removing missing values,
was split into training data (7,639 instances) and test data (955 instances).
A random forest model with 50 trees was trained using the training data.
Subsequently, LIME and R-LIME explanations were generated
for two instances extracted from the test data (\cref{fig:instance}).
For R-LIME, logistic regression was used as the linear approximation model,
and the distribution $\mathcal{D}$ was a multivariate normal distribution
estimated from the training data.
The beam width was set to $B=10$,
the confidence coefficient to $1-\delta=0.95$,
and the tolerance of the KL-LUCB algorithm to $\epsilon=0.05$.
Accuracy thresholds $\tau$ were set to $\tau=0.70,0.80,0.90$.

@@

\subsection{Changes in Reward Distribution in Best Arm Identification}\label{sec:reward}
{%
  \renewcommand{\arraystretch}{1.1}
  \begin{table}[tbp]
    \centering
    \caption[Deviation between the estimated accuracy and the true accuracy]{%
      Deviation between the estimated accuracy and the true accuracy.
      Deviation was relatively small considering confidence level $1-\delta=0.95$.
    }\label{tab:reward}
    \input{tex/tab_exp_adv}
  \end{table}
}
For R-LIME,
the problem of selecting the rule with the highest accuracy is formulated
as the optimal arm identification problem in multi-armed bandit theory,
solved using the KL-LUCB algorithm \cite{kaufmann2013information}.
However, this algorithm assumes that the reward distribution remains constant,
while in R-LIME,
the reward distribution (accuracy of the linear approximation)
changes with every update of the approximation model after sampling.
Therefore, rewards obtained at an early stage
might influence the estimated value and deviate from the true value.
